Investigating Trends of Absences in Toronto City Council Attendance between 2022 to 2026

### 2024-09-26 paper/paper.qmd
Updated LLM usage & some formating in paper

library(tidyverse)
library(dplyr)
library(readr)
library(ggplot2)

@@

# Introduction

@@

The paper is structured as follows: section 2 discusses the raw and clean data used in the study and any limitations in the dataset. Section 3 shows a graphical visualization of the data from section 2 and explains some trends that can be seen. Finally, section 4 covers interpretations of the data from section 3 and any weaknesses of this study.

@@

```{r echo=FALSE}
#| label: tbl-raw-data
head(raw_data, 10)

```

@@

After importing the data into R (@citeR) using the Open Data Toronto package (@OpenDataToronto), it is cleaned to keep only key variables of interest and make it easier for data analysis in Section 3. To clean the data and produce graphs with the data, the following packages in R (@citeR) were used: tidyverse (@tidyverse), dplyr (@dplyr), lubridate (@lubridate), knitr (@knitr), readr (@readr) and ggplot2 (@ggplot2). ChatGPT (@ChatGPT) was used to help the coding of cleaning, testing, and graphing data.

```{r echo=FALSE}
#| label: tbl-cleaned-data
head(cleaned_data, 10)

```
The key variables of interest are name, committee, session year, session type, and present. First and last names of City Council members are combined into 1 variable to make it easier to read in graphs. Then, the session date is decoded to separate the year, month and date - keeping only the year. When cleaning the data, only the columns of interest were selected to be in the newly cleaned dataset. The cleaned dataset consists of 5 variables as seen in @tbl-cleaned-data. Tests are run on the cleaned dataset to ensure there are no empty rows & that the remaining observations consist of only City Council members’ data.

@@

# Results

@@

```{r}
#| label: fig-session-member
#| fig-cap: Comparision of Absences by name
#| echo: false

@@

ggplot(absence_by_member, aes(x = reorder(name, -absence_count), y = absence_count)) +
  geom_bar(stat = "identity", fill = "cyan3") +
  labs(
    title = "Number of Absences by Each Member",
    x = "Council Member",
    y = "Number of Absences"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@@

To further our understanding of the data, @fig-session-year compares the number of member absences over the years since the beginning of the term. The x-axis is the years since the beginning of the term and the y-axis counts the number of absences.

```{r}
#| label: fig-session-year
#| fig-cap: Comparison of Member Absences by Year
#| echo: false

@@

# Create the plot
ggplot(absent_data, aes(x = session_year, y = absence_count)) +
  geom_bar(stat = "identity", fill = "cyan3") +  # Use a bar plot and custom color
  labs(
    title = "# absence over years 2022-2024",
    x = "Year",
    y = "# Absences"
  ) +
  theme_minimal()
```

@@

```{r}
#| label: fig-session-type
#| fig-cap: Comparison of Attendance by session type between 2022 to 2024
#| echo: false

@@

# Create the plot
ggplot(absence_data, aes(x = session_type, y = absence_count)) +
  geom_bar(stat = "identity", fill = "cyan3") +  # Use horizontal bar plot and a custom color
  labs(
    title = "# absence of different mettings types",
    x = "Type",
    y = "# Absences"
  ) +
  theme_minimal()
```

# Discussion

@fig-session-member from Section 3 laid out the information about the absences of each Toronto City Council member. It can be seen that Michael Thompson has the highest absences, followed by Jaye Robinson and Vincent Crisanti. While Amber Morley, Olivia Chow, and Paula Fletcher have missed few meetings. The variation between members could be problematic especially when the council member fails to represent the population. Additionally, it could suggest differences in engagement, personal schedules, or emergency circumstances. Determining those with the most absences allows steps to be taken to help better accommodate their schedules.

>>>>

library(tidyverse)
library(dplyr)
library(readr)
library(ggplot2)
library(knitr)

@@

# Introduction {#sec-intro}

@@

The paper is structured as follows: [@sec-data] discusses the raw and clean data used in the study and any limitations in the dataset. [@sec-results] shows a graphical visualization of the data from [@sec-data] and explains some trends that can be seen. Finally, [@sec-discussion] covers interpretations of the data from [@sec-results] and any weaknesses of this study.

@@

```{r echo=FALSE}
#| label: tbl-raw-data
#| tbl-cap: Raw Data of Toronto City Council Members' Attendance

# generate first 10 observations of raw data
kable(head(raw_data, 10)) 
```

@@

After importing the data into R (@citeR) using the Open Data Toronto package (@OpenDataToronto), it is cleaned to keep only key variables of interest and make it easier for data analysis in [@sec-results]. To clean the data and produce graphs with the data, the following packages in R (@citeR) were used: tidyverse (@tidyverse), dplyr (@dplyr), lubridate (@lubridate), knitr (@knitr), readr (@readr) and ggplot2 (@ggplot2). ChatGPT (@ChatGPT) was used to help the coding of cleaning, testing, and graphing data.

```{r echo=FALSE}
#| label: tbl-cleaned-data
#| tbl-cap: Cleaned Data of Toronto City Council Members' Attendance

# generate first 10 observations of cleaned data
kable(head(raw_data, 10)) 
```
The key variables of interest are name, committee, session year, session type, and present. First and last names of City Council members are combined into 1 variable to make it easier to read in graphs. Then, the session date is decoded to separate the year, month and date - keeping only the year. When cleaning the data, only the columns of interest were selected to be in the newly cleaned dataset. The cleaned dataset consists of 5 variables as seen in @tbl-cleaned-data. Tests are run on the cleaned dataset to ensure there are no empty rows & that the remaining observations consist of only City Council members’ data.

@@

# Results {#sec-results}

@@

```{r}
#| label: fig-session-member
#| fig-cap: Comparision of Absences by City Council Member
#| echo: false

@@

# generate graph for number of absences by each member
ggplot(absence_by_member, aes(x = reorder(name, -absence_count), y = absence_count)) +
  geom_bar(stat = "identity", fill = "cyan3") +
  labs(
    title = "Number of Absences by Each Member",
    x = "Council Member",
    y = "Number of Absences"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@@

To further our understanding of the data, @fig-session-year compares the number of member absences over the years since the beginning of the term. The x-axis is the years since the beginning of the term until the present year and the y-axis counts the number of absences.

```{r}
#| label: fig-session-year
#| fig-cap: Comparison of Absences by Year
#| echo: false

@@

# Create the plot
ggplot(absent_data, aes(x = session_year, y = absence_count)) +
  geom_bar(stat = "identity", fill = "cyan3") +  # Use a bar plot and custom color
  labs(
    title = "Number of Absence Over Years 2022-2024",
    x = "Year",
    y = "Number of Absences"
  ) +
  theme_minimal()
```

@@

```{r}
#| label: fig-session-type
#| fig-cap: Comparison of Absences by Session Type
#| echo: false

@@

# Create the plot
ggplot(absence_data, aes(x = session_type, y = absence_count)) +
  geom_bar(stat = "identity", fill = "cyan3") +  # Use horizontal bar plot and a custom color
  labs(
    title = "Number Absence for Different Meetings Types",
    x = "Session Type",
    y = "Number of Absences"
  ) +
  theme_minimal()
```

# Discussion {#sec-discussion}

@fig-session-member from [@sec-results] laid out the information about the absences of each Toronto City Council member. It can be seen that Michael Thompson has the highest absences, followed by Jaye Robinson and Vincent Crisanti. While Amber Morley, Olivia Chow, and Paula Fletcher have missed few meetings. The variation between members could be problematic especially when the council member fails to represent the population. Additionally, it could suggest differences in engagement, personal schedules, or emergency circumstances. Determining those with the most absences allows steps to be taken to help better accommodate their schedules.
